---
title: "Data Processing Phase 2 Report"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Data Processing Phase 2 Report}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This report focuses on the data processing steps involved in phase 2. It
encompasses filtering the data, transforming it using Compositional
Log-Ratio (CLR) transformation, assessing batch effects, and applying
batch correction methods such as removeBatchEffect, ComBat, PLSDA-batch,
sPLSDA-batch, and percentile normalization.

## Setup

In this section, we set up the necessary libraries and configurations
for the data processing tasks.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
# Load necessary libraries
library(dataPreparation)
library(PLSDAbatch)
library(mixOmics)
library(patchwork)
library(vegan)
library(UpSetR)
```

## Data Loading

```{r}
categories <- dataPreparation::imputed_data$categories
imputed_data <- dataPreparation::imputed_data
```

# Preparing Data for Analysis

We begin by filtering the data using the `PreFL()` function from the
PLSDAbatch package to remove features with zero variance. This step is
crucial for reducing noise and improving the quality of the dataset.

```{r}
# Filter the data
filter.res <- PLSDAbatch::PreFL(data = imputed_data$data, 
                                keep.spl = 0, 
                                keep.var = 0.00)
filter <- filter.res$data.filter

# Calculate zero proportion before filtering
filter.res$zero.prob

# Calculate zero proportion after filtering
sum(filter == 0) / (nrow(filter) * ncol(filter))
```

# Transforming Data

Next, I transform the filtered data using the CLR transformation method
from the mixOmics package. This transformation is essential for handling
compositional data and preparing it for further analysis.

```{r}
# Perform CLR transformation
clr <- mixOmics::logratio.transfo(X = filter, logratio = 'CLR', offset = 1) 
class(clr) = 'matrix'
```

# Assessing Batch Effects

Before applying batch correction methods, we assess batch effects in the
data using principal component analysis (PCA) and variance partitioning
analysis (pRDA). Understanding the sources of variation in the data is
crucial for selecting appropriate batch correction techniques.

```{r}
# Perform PCA
pca.before <- mixOmics::pca(clr, ncomp = 4, scale = TRUE)

batch = factor(categories$Submission.Name, levels = unique(categories$Submission.Name))

descr = as.factor(categories$Sample.Description)

names(batch) <- names(descr) <- rownames(categories)

# Perform pRDA
factors.df <- data.frame(trt = descr, batch = batch)
rda.before <- vegan::varpart(clr, ~ descr, ~ batch, 
                             data = factors.df, 
                             scale = TRUE)
rda.before$part$indfract
```

# Batch Correction

I apply various batch correction methods, including `removeBatchEffect`,
`ComBat`, `PLSDA-batch`, and `sPLSDA-batch`, to mitigate batch effects
in the data. These methods adjust for technical variation introduced by
batch processing and improve the accuracy of downstream analysis.

```{r}
# Managing batch effects
clr <- clr[seq_len(nrow(clr)), seq_len(ncol(clr))]
lm <- PLSDAbatch::linear_regres(data = clr, trt = descr, 
                       batch.fix = batch, type = 'linear model')

p <- sapply(lm$lm.table, function(x){x$coefficients[2,4]})
p.adj <- p.adjust(p = p, method = 'fdr')

performance::check_model(lm$model$AC.0.0.)

mod <- model.matrix( ~ descr)
```

```{r}
# Applying removeBatchEffect
rBE <- t(limma::removeBatchEffect(t(clr), batch = batch, 
                              design = mod))

# Applying ComBat
ComBat <- t(sva::ComBat(t(clr), batch = batch, 
                      mod = mod, par.prior = FALSE))

# Applying PLSDA-batch
trt.tune <- mixOmics::plsda(X = clr, Y = descr, ncomp = 5)
trt.tune$prop_expl_var #1
ad.batch.tune <- PLSDAbatch::PLSDA_batch(X = clr, 
                             Y.trt = descr, Y.bat = batch,
                             ncomp.trt = 1, ncomp.bat = 10)
ad.batch.tune$explained_variance.bat #4
PLSDA_batch.res <- PLSDAbatch::PLSDA_batch(X = clr, 
                                  Y.trt = descr, Y.bat = batch,
                                  ncomp.trt = 1, ncomp.bat = 4)
PLSDA_batch <- PLSDA_batch.res$X.nobatch

# Applying sPLSDA-path
set.seed(777)
test.keepX = c(seq(1, 10, 1), seq(20, 100, 10), 
                  seq(150, 231, 50), 231)
trt.tune.v <- mixOmics::tune.splsda(X = clr, Y = descr, 
                             ncomp = 1, test.keepX = test.keepX, 
                             validation = 'Mfold', folds = 4, 
                             nrepeat = 50)
trt.tune.v$choice.keepX
batch.tune <- PLSDAbatch::PLSDA_batch(X = clr, 
                             Y.trt = descr, Y.bat = batch,
                             ncomp.trt = 1, keepX.trt = 100,
                             ncomp.bat = 10)
batch.tune$explained_variance.bat #4
sum(batch.tune$explained_variance.bat$Y[seq_len(4)])
sPLSDA_batch.res <- PLSDAbatch::PLSDA_batch(X = clr, 
                                   Y.trt = descr, Y.bat = batch,
                                   ncomp.trt = 1, keepX.trt = 100,
                                   ncomp.bat = 4)
sPLSDA_batch <- sPLSDA_batch.res$X.nobatch

# Applying PN
PN <- PLSDAbatch::percentile_norm(data = clr, batch = batch, 
                         trt = descr, ctrl.grp = '0-0.5')

```

```{r eval=FALSE, include=FALSE}
# # RUVIII
# replicates <- categories |>
#   colnames()
# 
# # empirical negative controls
# empir.p <- c()
# for(e in seq_len(ncol(clr))){
#   empir.lm <- lm(clr[,e] ~ descr)
#   empir.p[e] <- summary(empir.lm)$coefficients[2,4]
# }
# empir.p.adj <- p.adjust(p = empir.p, method = 'fdr')
# nc <- empir.p.adj > 0.05
# # estimate k
# # k.res <- kyotil::getK(Y = clr, X = descr, ctl = nc)
# k.res <- kyotil::getK(X = descr, kernel = "ibs", X2 = clr)
# k <- k.res$k
# 
# replicates.matrix <- ruv::replicate.matrix(replicates)
# 
# RUVIII <- ruv::RUVIII(Y = clr, M = replicates.matrix,
#                     ctl = nc, k = k)
# rownames(RUVIII) <- rownames(clr)
```

# Evaluating Batch Correction

To evaluate the effectiveness of batch correction methods, I compare the
variance explained by treatment and batch factors before and after
correction. I also assess the impact of correction on the distribution
of samples using scatter plots and density plots.

```{r}
# Perform PCA after batch correction
pca.rBE <- mixOmics::pca(rBE, ncomp = 3, scale = TRUE)
pca.ComBat <- mixOmics::pca(ComBat, ncomp = 3, scale = TRUE)
pca.PLSDA_batch <- mixOmics::pca(PLSDA_batch, ncomp = 3, scale = TRUE)
pca.sPLSDA_batch <- mixOmics::pca(sPLSDA_batch, ncomp = 3, scale = TRUE)
```

# Selecting Features

Finally, we select features that are relevant for discrimination between
treatment groups using sparse partial least squares discriminant
analysis (sPLSDA). This step helps identify biomarkers or features that
contribute significantly to group separation and biological
interpretation.

```{r}
# order batches
batch <- factor(categories$Submission.Name, 
               levels = unique(categories$Submission.Name))

pca.before.plot <- PLSDAbatch::Scatter_Density(object = pca.before, 
                                      batch = batch, 
                                      trt = descr, 
                                      title = 'Before correction')
pca.rBE.plot <- PLSDAbatch::Scatter_Density(object = pca.rBE, 
                                   batch = batch, 
                                   trt = descr, 
                                   title = 'removeBatchEffect')
pca.ComBat.plot <- PLSDAbatch::Scatter_Density(object = pca.ComBat, 
                                      batch = batch, 
                                      trt = descr, 
                                      title = 'ComBat')
pca.PLSDA_batch.plot <- PLSDAbatch::Scatter_Density(object = pca.PLSDA_batch, 
                                           batch = batch, 
                                           trt = descr, 
                                           title = 'PLSDA-batch')
pca.sPLSDA_batch.plot <- PLSDAbatch::Scatter_Density(object = pca.sPLSDA_batch, 
                                            batch = batch, 
                                            trt = descr, 
                                            title = 'sPLSDA-batch')

g <- ggpubr::ggarrange(pca.before.plot, 
          pca.rBE.plot,
          pca.ComBat.plot,
          pca.PLSDA_batch.plot,
          pca.sPLSDA_batch.plot,
          labels = c("A", "B", "C", "D", "E"),
          ncol = 2, nrow = 3)
```

```{r include=FALSE}
grDevices::png("../man/figures/batch_correction1.png", units="px", width=2100, height=2100, res=210)
print(g)
dev.off()
```

```{r}
corrected.list <- list(`Before correction` = clr, 
                          removeBatchEffect = rBE, 
                          ComBat = ComBat,
                          `PLSDA-batch` = PLSDA_batch,
                          `sPLSDA-batch` = sPLSDA_batch
                          # `Percentile Normalisation` = PN,
                          # RUVIII = RUVIII
                          )
```

```{r}
factors.df <- data.frame(trt = descr, batch = batch)

prop.df <- data.frame("Disease Pheno" = NA, Batch = NA, 
                         Intersection = NA, 
                         Residuals = NA) 
for(i in seq_len(length(corrected.list))){
  rda.res = vegan::varpart(corrected.list[[i]], ~ descr, ~ batch,
                    data = factors.df, scale = TRUE)
  prop.df[i, ] <- rda.res$part$indfract$Adj.R.squared}

rownames(prop.df) = names(corrected.list)

prop.df <- prop.df[, c(1,3,2,4)]

prop.df[prop.df < 0] = 0
prop.df <- as.data.frame(t(apply(prop.df, 1, 
                                    function(x){x/sum(x)})))

PLSDAbatch::partVar_plot(prop.df = prop.df)

```

## other methods

```{r}
d <-
  dataPreparation::visualize_batch_correction(
    corrected_list = corrected.list,
    categories = categories,
    visualization_type = "barplot"
  )
d
```

```{r include=FALSE}
png("../man/figures/batch_correction2.png", units="px", width=2200, height=1100, res=300)
print(d)
dev.off()
```



```{r}
splsda.select <- list()
for(i in seq_len(length(corrected.list))){
  splsda.res <- mixOmics::splsda(X = corrected.list[[i]], Y = descr, 
                       ncomp = 3, keepX = rep(50,3))
  select.res <- mixOmics::selectVar(splsda.res, comp = 1)$name
  splsda.select[[i]] <- select.res
}
names(splsda.select) <- names(corrected.list)

# can only visualize 5 methods
splsda.select <- splsda.select[seq_len(5)]

splsda.upsetR <- UpSetR::fromList(splsda.select)

p <-  UpSetR::upset(splsda.upsetR, main.bar.color = 'gray36',
      sets.bar.color = PLSDAbatch::pb_color(c(25:22,20)), matrix.color = 'gray36',
      order.by = 'freq', empty.intersections = 'on',
      queries = list(list(query = intersects, 
                          params = list('Before correction'), 
                          color = PLSDAbatch::pb_color(20), active = TRUE),
                     list(query = intersects, 
                          params = list('removeBatchEffect'), 
                          color = PLSDAbatch::pb_color(22), active = TRUE),
                     list(query = intersects, 
                          params = list('ComBat'), 
                          color = PLSDAbatch::pb_color(23), active = TRUE),
                     list(query = intersects, 
                          params = list('PLSDA-batch'), 
                          color = PLSDAbatch::pb_color(24), active = TRUE),
                     list(query = intersects, 
                          params = list('sPLSDA-batch'), 
                          color = PLSDAbatch::pb_color(25), active = TRUE)))

```

# Display Up

```{r}
p
```

```{r include=FALSE}
png("../man/figures/splsda_upsetR.png")
print(p)
dev.off()
```
# Performing Principal Component Analysis (PCA After Batch Effect Correction) 

Principal Component Analysis (PCA) is a dimensionality reduction technique
that helps visualize the variation in the dataset. Here, I perform PCA on
the preprocessed data and create a PCA plot for visualization.

```{r}
t <- rBE
# Perform PCA
pca_result <- prcomp(t, scale. = TRUE, center = TRUE)

# Extract PCA scores
pca_scores <- as.data.frame(pca_result$x)

# Combine PCA scores with categories for visualization
pca_data <- cbind(pca_scores, imputed_data$categories)

# Create PCA plot
pca_plot <- ggplot2::ggplot(pca_data, 
                            ggplot2::aes(x = PC1, y = PC2, 
                                         color = Submission.Name, 
                                         shape = Sample.Description)) +
    ggplot2::geom_point(size = 2) +
    ggplot2::scale_shape_manual(values = c(16, 17, 18, 19)) +
    ggplot2::theme_bw() +
    ggplot2::labs(
      title = "PCA Plot (After Batch Effect Correction)",
      x = paste0(
        "PC1 (Explained Variance: ",
        round(pca_result$sdev[1] ^ 2 / sum(pca_result$sdev ^ 2) * 100, 2),
        "%)"
      ),
      y = paste0(
        "PC2 (Explained Variance: ",
        round(pca_result$sdev[2] ^ 2 / sum(pca_result$sdev ^ 2) * 100, 2),
        "%)"
      )
    )

# Display PCA plot
pca_plot
```


```{r}
imputed_data$rBE <- rBE
write.csv(x = imputed_data$rBE, 
          file = "../inst/data_to_use/imputed_data_after_Batch_Correction.csv", 
          row.names = FALSE)
usethis::use_data(imputed_data, overwrite = TRUE)
```


```{r}
png("../man/figures/pca_plot_after_Batch_Correction.png", 
    width=1600, height=1200, res=220)
print(pca_plot)
dev.off() 
```


# Conclusion

In conclusion, this report highlights the importance of comprehensive
data processing and batch correction techniques in ensuring the
reliability and interpretability of metabolomic data. By systematically
addressing batch effects and selecting informative features, I can
improve the robustness f downstream analyses and enhance the
understanding of biological phenomena.

This marks the end of the Data Processing Phase 2 Report.

# Clean up environment

```{r}
# rm(list = ls())
```


